A general and efficient representation of Ancestral Recombination Graphs

### 2023-08-09 paper.tex

Recent breakthroughs in large-scale inference
methods~\citep{rasmussen2014genome,kelleher2019inferring,speidel2019method,wohns2022unified}
and data representation~\citep{kelleher2016efficient}
have raised the realistic prospect of genealogy-based analysis becoming a standard part
of the population and statistical genetics toolkit~\citep{hejase2020summary}.
Applications using these inferred ancestries as inputs have begun to
appear~\citep{osmond2021estimating,zhang2021biobank,fan2022genealogical,hejase2022deep}
and many more are sure to
follow~\citep{harris2019database}. This vibrant research area, however,
has a significant difficulty, which risks hampering
progress: we currently lack a well-defined data-model and shared terminology
to discuss genealogies in the presence of recombination. This leads to basic errors in statements
about what different inference methods produce as output, as well as presenting serious
difficulties in either comparing the outputs of different methods or
(from a user's perspective) using multiple inference methods in an analysis.

@@

\section*{Precision of ancestry information}
% Not all nodes in an ARG are equal. Assuming that we have sequenced the
% genomes of the samples, there are intrinsic limits on what can
% be known about more ancient nodes, determined by the graph topology
% and the flow of ancestral material through it.

Ancestral recombination graphs are often characterised
as consisting of a complete record of the genetic history
of a sample. For example,
\cite{rasmussen2014genome} state that an ``ARG provides a record of all
coalescence and recombination events since the divergence of the sequences
under study'' and
according to \cite{deng2021distribution} an ARG
``provides all the information about the genealogical history of a sample,
including the locations of recombination events.'' It is worth
questioning, however, whether such comprehensive information
will always be available, and in particular whether our
representation of genealogical history should \emph{require}
such precision. Simulations will usually generate complete
information about the history of a sample, of course, but
we may not have sufficient information to infer such detail.

% Not sure where this goes, but there are some critical points
% that we want to explain and not just throw out there in the
% discussion.
A critical distinction between the eARG and gARG encodings
is the level of precision about the time and location of
recombination events that is required. In an eARG, we are
required to be absolutely precise about every recombination
[explain and contrast with gARG, referring to examples]

% % JK: This is very rough, just jotting down the basic ideas,
% % will need substantial revision.
% It is not clear that such complete information will always
% (or indeed ever) be available from inferences. We may be
% able to precisely identify the location and timing
% of recent recombinations (particularly if detailed pedigree
% information is available) but beyond this, the information
% required simply isn't present. When sampling from a
% process such as the Sequentially Markov
% Coalescent~\citep{mcvean2005approximating,marjoram2006fast}
% like ARGweaver~\citep{rasmussen2014genome}, it is true that
% we do obtain a fully resolved ARG with precise
% information about recombination, but it is reasonable
% to question how much information from the data is used
% to support older recombinations, in particular.

% It is helpful to consider the difference between fully
% resolved binary trees and those that contain polytomies.

% The situation is in fact directly analogous: an ARG that
% contains fully resolved recombination information must have
% at most parents per node.

% At the very least, it is not reasonable for the \emph{data
% structure} to require complete precision in order to
% represent some less precise. This is a fundamental
% advantage of the gARG encoding over the classical eARG
% method: we are not forced to posit a precise cause (recombination
% event) for every effect (difference between adjacent trees).
% These differences in precision of recombination
% information are illustrated in the ARGs estimated by different
% methods, as we see in the next section.

To illustrate this critical point, we inferred ARGs for the
the classical \citet{kreitman1983nucleotide} benchmark dataset
using four recent inference methods and show the results in
Fig~\ref{fig-inferred-args}.
While there is some consistency among the inferred ARGs (such as the samples
FR-F, WA-F and Af-F being grouped into the same clade), there
are also considerable differences. For this dataset, the minimum number
of recombination breakpoints required in the absence of recurrent mutation
is 7~\citep{song2003parsimonious}, and the number of inferred breakpoints
under the input mutation and recombination rates
varies from 2 (\relate) to 17 (\argweaver).
The ARGs inferred by the different methods also have major differences
in how recombination information is represented. Essentially,
\kwarg\ and \argweaver\ infer a specific causal recombination \emph{event}
for each breakpoint, and those events are fully
described in the output graph. On the other hand, \tsinfer\ and
\relate\ do not infer causal events but instead focus on
representing the inferred \emph{effect}: breakpoints between the
trees. It is important to note that this does not imply that
the methods must then necessarily infer only a sequence of
unrelated local trees
% More? See https://github.com/tskit-dev/what-is-an-arg-paper/issues/38
% for discussion. I'm sure there were other examples of people saying
% that
(as claimed by e.g.~\citet{hejase2020summary}),
and information about recombination and the correlation structure
between trees is not ``all or nothing''.
Both methods capture correlation structure between trees:
\relate\ first infers local tree topologies and then identifies edges
that persist from one tree to the next as a separate step
(in a similar manner to \espalier~\citep{rasmussen2022espalier}), while
\texttt{tsinfer} explicitly constructs nodes that are shared across local
trees.

\section*{When do we need ``full'' ARGs?}
% The full ancestral recombination graph (ARG) is a structure that encodes all
% coalescence and recombination events resulting from the stochastic process of
% the coalescent with recombination.
\citet{brandt2021evaluation} define a ``full'' ARG as
``a structure that encodes all
coalescence and recombination events resulting from the stochastic process of
the coalescent with recombination'' (thus going against the general
historial trend discussed in the XXX section).
\citet{rasmussen2022espalier} have a less restrictive definition, in
which a ``full'' ARG is one that contains complete information about
recombination events, without necessarily being tied to a particular
stochastic process.
While [something positive], it is important to be clear on
what the overall goal of such precise inferences are,
what the limitations on when they can be made are,
and their downstream utility in applications.

@@

% If we want to use the coalescent, then really the assumptions
% need to make sense.
% n << Ne is a silly assumption in today's datasets
A core assumption of the coalescent is that the sample size $n$
is much less than the effective population size, $N_e$.
Several human datasets now consist of hundreds of thousands of
genomes~\citep{bycroft2018genome,karczewski2020mutational,tanjo2021practical},
and so sample size is substantially \emph{larger} than $N_e$
(often assumed to be $10^4$ in humans).
Agricultural datasets are even more extreme.
For example,
the US dairy cattle database alone currently comprises more than 6 million
animals with SNP array
genotypes\footnote{\url{https://queries.uscdcb.com/Genotype/counts.html}},
while the effective population size in modern dairy cattle breeds is
less than 100 and decreasing~\citep{MacLeod2013,Makanjuola2020}.
An extreme sign of these breeding practices is that there are only two ancestral
Y-chromosome lineages present in today's US Holstein dairy breed~\citep{Yue2015}.
Simlarly, the 1000 bull genomes project~\citep{hayes20191000}
comprises close to 7000 genomes, which are part of multi-generation pedigrees
with millions of animals and extensive SNP array genotype and phenotype
data \citep[e.g.][]{Cesarani2022}.
In another example, genome and SNP array data for 440,610 individuals within
7 multi-generation pedigrees~\citep{whalen2018,Johnsson2021,Ros-Freixedes2020}
were combined to infer recombination in pigs~\citep{RosFreixedes2022}.
Effective population size in modern pig breeds is also less than 100 due to
intense selection and directed reproduction \citep{Hall2016,Porcnic2016}.
Another core assumption of the coalescent model is that the genome (or
at least the region under study) is short enough that the number of extant
lineages remains much smaller than $N_e$ at all times. Whole
genome sequences have been available for model organisms
for over a decade now, % True? Citation?
and indeed complete chromosome-level assemblies are possible
in humans~\citep{miga2020telomere}.
Projects are under way to obtain high-quality assemblies
for all eukaryotic species in Britain and Ireland~\citep{darwin2022sequence}
and ultimately worldwide~\citep{lewin2022earth}.

While the coalescent as a model can be suprisingly robust to
basic violations of the underlying assumptions~\citep{
wakeley2012gene,bhaskar2014distortion,nelson2020accounting}
it is reasonable to question the validity of the model in
representing datasets such as UK Biobank.
[Also mention somehow that $n\ll N_e$ is only one problem -
we're also ignoring the myriad other evolutionary forces.]
Such datasets cannot be reasonably be described by
the coalescent as a statical model, and so more non-parametric
methods like \tsinfer\ and \relate\ would seem more appropriate.
[More]

% Is it practical to estimate full ARGs?
% [Not particularly \citep{shipilina2022origin} - but maybe ARGneedle
% has done this for UKB? Need to check]

% Do downstream applications actually make use of such full ARGs?
Downstream applications using estimated ARGs to date rarely
make use of such complete recombination information.
The \relate\ selection statistic~\citep{speidel2019method}
computes a $p$-value for selection [FIXME] for each
local tree independently by comparing the rates of coalescence
at various time-points. For cases like strong positive selection,
where we expect there to be a clear signal in the local tree,
such tree-by-tree methods can be simple and powerful.
In their method
for estimating dispersal rates and the locations of genetic
ancestors,
\citep{osmond2021estimating} downsample trees along the genome
so that they can be regarded as approximately independent.
The SIA method for detecting selection~\citep{hejase2022deep}
encodes local trees as a set of lineage counts at discrete
time intervals, and uses these as feature for a
type of machine learning algorithm
that takes ``temporal'' correlations into account. Thus,
while SIA does use some of the information about local tree correlation,
clearly much of the detail about recombination
events in a ``full'' ARG is lost.
The \texttt{tsdate} algorithm (and related naive estimator
for ancestral location) uses much more between-tree
information~\citep{wohns2022unified}, explicitly using the gARG
encoding to reason about ancestral nodes.
A major strength of this approach is that it can be used on
any gARG, ranging from estimates containing relatively little
information about recombination (e.g., as estimated by \relate)
to ``full'' ARGs containing complete recombination information
(e.g., as estimated by \kwarg), and will fully utilise
any between-tree correlation information that is present.

\section*{Computational efficiency}
The succinct tree sequence data structure (usually known as a ``tree sequence''
for brevity, but see later comment on confusion around this point)
is a concrete encoding of a gARG, as discussed in this article.
It was originally developed as part of the \texttt{msprime}
simulator~\cite{kelleher2016efficient} and subsequently ...

[The tree sequence encoding is very efficient for lots of things. Examples]

In retrospect, the choice of terminology was unfortunate, and has led to
significant confusion. [More]

@@

Reconstructing a most parsimonious ARG for a given data set is NP-hard
\citep{wang2001perfect}, so parsimony-based methods resort to heuristics and
are limited to analysing at most hundreds of sequences. Hence, a number of
methods aim to balance computational efficiency with reconstruction of
``reasonable", rather than parsimonious ARGs \citep{minichiello2006mapping,
parida2008estimating, kelleher2019inferring,  speidel2019method,
zhang2021biobank}. The latter three methods, as well as the parsimony-based
method of \citet{rasmussen2022espalier}, can be used on megabase scale data and
hundreds of thousands of samples under human-like parameters.

>>>>

Recent breakthroughs in large-scale inference
methods~\citep{rasmussen2014genome,kelleher2019inferring,speidel2019method,wohns2022unified}
and data representation~\citep{kelleher2016efficient}
have raised the realistic prospect of genealogy-based analysis becoming a standard part
of the population and statistical genetics toolkit~\citep{hejase2020summary}.
Applications using these inferred ancestries as inputs have begun to
appear~\citep{osmond2021estimating,zhang2023biobank,fan2022genealogical,hejase2022deep}
and many more are sure to
follow~\citep{harris2019database}. This vibrant research area, however,
has a significant difficulty, which risks hampering
progress: we currently lack a well-defined data-model and shared terminology
to discuss genealogies in the presence of recombination. This leads to basic errors in statements
about what different inference methods produce as output, as well as presenting serious
difficulties in either comparing the outputs of different methods or
(from a user's perspective) using multiple inference methods in an analysis.

@@

% NEEDS more work!
\section*{Precision of ancestry information}
Ancestral recombination graphs are often characterised
as consisting of a complete record of the genetic history
of a sample. For example,
\cite{rasmussen2014genome} state that an ``ARG provides a record of all
coalescence and recombination events since the divergence of the sequences
under study'' and
according to \cite{deng2021distribution} an ARG
``provides all the information about the genealogical history of a sample,
including the locations of recombination events.'' It is worth
questioning, however, whether such comprehensive information
will always be available, and in particular whether our
representation of genealogical history should \emph{require}
such precision. Simulations will usually generate complete
information about the history of a sample, of course, but
we may not have sufficient information to infer such detail.

% The full ancestral recombination graph (ARG) is a structure that encodes all
% coalescence and recombination events resulting from the stochastic process of
% the coalescent with recombination.
\citet{brandt2021evaluation} define a ``full'' ARG as
``a structure that encodes all
coalescence and recombination events resulting from the stochastic process of
the coalescent with recombination'' (thus going against the general
historial trend discussed in the XXX section).
\citet{rasmussen2022espalier} have a less restrictive definition, in
which a ``full'' ARG is one that contains complete information about
recombination events, without necessarily being tied to a particular
stochastic process.
While [something positive], it is important to be clear on
what the overall goal of such precise inferences are,
what the limitations on when they can be made are,
and their downstream utility in applications.

@@

% Not sure where this goes, but there are some critical points
% that we want to explain and not just throw out there in the
% discussion.
% A critical distinction between the eARG and gARG encodings
% is the level of precision about the time and location of
% recombination events that is required. In an eARG, we are
% required to be absolutely precise about every recombination
% [explain and contrast with gARG, referring to examples]

% JK: This is very rough, just jotting down the basic ideas,
% will need substantial revision.
It is not clear that such complete information will always
(or indeed ever) be available from inferences. We may be
able to precisely identify the location and timing
of recent recombinations (particularly if detailed pedigree
information is available) but beyond this, the information
required simply isn't present. When sampling from a
process such as the Sequentially Markov
Coalescent~\citep{mcvean2005approximating,marjoram2006fast}
like ARGweaver~\citep{rasmussen2014genome}, it is true that
we do obtain a fully resolved ARG with precise
information about recombination, but it is reasonable
to question how much information from the data is used
to support older recombinations, in particular.

It is helpful to consider the difference between fully
resolved binary trees and those that contain polytomies.

The situation is in fact directly analogous: an ARG that
contains fully resolved recombination information must have
at most parents per node.

At the very least, it is not reasonable for the \emph{data
structure} to require complete precision in order to
represent some less precise. This is a fundamental
advantage of the gARG encoding over the classical eARG
method: we are not forced to posit a precise cause (recombination
event) for every effect (difference between adjacent trees).
These differences in precision of recombination
information are illustrated in the ARGs estimated by different
methods, as we see in the next section.

To illustrate these points and to emphasise the heterogeneity
of the ARGs inferred by current methods, we inferred ARGs for the
the classical \citet{kreitman1983nucleotide} benchmark dataset
using four recent inference methods and show the results in
Fig~\ref{fig-inferred-args}.
While there is some consistency among the inferred ARGs (such as the samples
FR-F, WA-F and Af-F being grouped into the same clade), there
are also considerable differences. For this dataset, the minimum number
of recombination breakpoints required in the absence of recurrent mutation
is 7~\citep{song2003parsimonious}, and the number of inferred breakpoints
under the input mutation and recombination rates
varies from 2 (\relate) to 17 (\argweaver).
The ARGs inferred by the different methods also have major differences
in how recombination information is represented. Essentially,
\kwarg\ and \argweaver\ infer a specific causal recombination \emph{event}
for each breakpoint, and those events are fully
described in the output graph. On the other hand, \tsinfer\ and
\relate\ do not infer causal events but instead focus on
representing the inferred \emph{effect}: breakpoints between the
trees. It is important to note that this does not imply that
the methods must then necessarily infer only a sequence of
unrelated local trees
% More? See https://github.com/tskit-dev/what-is-an-arg-paper/issues/38
% for discussion. I'm sure there were other examples of people saying
% that
(as claimed by e.g.~\citet{hejase2020summary}),
and information about recombination and the correlation structure
between trees is not ``all or nothing''.
Both methods capture correlation structure between trees:
\relate\ first infers local tree topologies and then identifies edges
that persist from one tree to the next as a separate step
(in a similar manner to \espalier~\citep{rasmussen2022espalier}), while
\texttt{tsinfer} explicitly constructs nodes that are shared across local
trees. See \cite{ignatieva2023distribution} for a detailed discussion
and comparsion of inference methods on this property.

% \section*{When do we need ``full'' ARGs?}
% % The full ancestral recombination graph (ARG) is a structure that encodes all
% % coalescence and recombination events resulting from the stochastic process of
% % the coalescent with recombination.
% \citet{brandt2021evaluation} define a ``full'' ARG as
% ``a structure that encodes all
% coalescence and recombination events resulting from the stochastic process of
% the coalescent with recombination'' (thus going against the general
% historial trend discussed in the XXX section).
% \citet{rasmussen2022espalier} have a less restrictive definition, in
% which a ``full'' ARG is one that contains complete information about
% recombination events, without necessarily being tied to a particular
% stochastic process.
% While [something positive], it is important to be clear on
% what the overall goal of such precise inferences are,
% what the limitations on when they can be made are,
% and their downstream utility in applications.

% % What is the goal of having fully precise ARGs?
% % 1) studying the actual recombination events;
% % 2) evaluating the likelihoods
% The main reasons for wanting precise estimates of all recombination
% events are to either compute a likelihood of the inferred ARG
% under a statistical model, or to study the recombinants in
% detail~\citep{rasmussen2022espalier}.
% % Explain that we're really interested in the details of recombination
% % itself with a small number of samples.
% In the latter case, we might
% have a small number of samples~\citep{guo2022recombination}.

% Beyond cases in which visual inspection of the trees and the
% effects of recombination is feasible, the main application
% of having completely precise estimates of recombination events is
% to evaluate the likelihood of the estimate under some
% statistical model. The only statistical model available
% is the coalescent with recombination
% and its approximations~\citep{mcvean2005approximating,marjoram2006fast}.
% In this case, the utility of a ``full`` ARG for some dataset
% is limited by how well the dataset fits the assumptions
% of the model.

% % If we want to use the coalescent, then really the assumptions
% % need to make sense.
% % n << Ne is a silly assumption in today's datasets
% A core assumption of the coalescent is that the sample size $n$
% is much less than the effective population size, $N_e$.
% Several human datasets now consist of hundreds of thousands of
% genomes~\citep{bycroft2018genome,karczewski2020mutational,tanjo2021practical},
% and so sample size is substantially \emph{larger} than $N_e$
% (often assumed to be $10^4$ in humans).
% Agricultural datasets are even more extreme.
% For example,
% the US dairy cattle database alone currently comprises more than 6 million
% animals with SNP array
% genotypes\footnote{\url{https://queries.uscdcb.com/Genotype/counts.html}},
% while the effective population size in modern dairy cattle breeds is
% less than 100 and decreasing~\citep{MacLeod2013,Makanjuola2020}.
% An extreme sign of these breeding practices is that there are only two ancestral
% Y-chromosome lineages present in today's US Holstein dairy breed~\citep{Yue2015}.
% Simlarly, the 1000 bull genomes project~\citep{hayes20191000}
% comprises close to 7000 genomes, which are part of multi-generation pedigrees
% with millions of animals and extensive SNP array genotype and phenotype
% data \citep[e.g.][]{Cesarani2022}.
% In another example, genome and SNP array data for 440,610 individuals within
% 7 multi-generation pedigrees~\citep{whalen2018,Johnsson2021,Ros-Freixedes2020}
% were combined to infer recombination in pigs~\citep{RosFreixedes2022}.
% Effective population size in modern pig breeds is also less than 100 due to
% intense selection and directed reproduction \citep{Hall2016,Porcnic2016}.
% Another core assumption of the coalescent model is that the genome (or
% at least the region under study) is short enough that the number of extant
% lineages remains much smaller than $N_e$ at all times. Whole
% genome sequences have been available for model organisms
% for over a decade now, % True? Citation?
% and indeed complete chromosome-level assemblies are possible
% in humans~\citep{miga2020telomere}.
% Projects are under way to obtain high-quality assemblies
% for all eukaryotic species in Britain and Ireland~\citep{darwin2022sequence}
% and ultimately worldwide~\citep{lewin2022earth}.

% While the coalescent as a model can be suprisingly robust to
% basic violations of the underlying assumptions~\citep{
% wakeley2012gene,bhaskar2014distortion,nelson2020accounting}
% it is reasonable to question the validity of the model in
% representing datasets such as UK Biobank.
% [Also mention somehow that $n\ll N_e$ is only one problem -
% we're also ignoring the myriad other evolutionary forces.]
% Such datasets cannot be reasonably be described by
% the coalescent as a statical model, and so more non-parametric
% methods like \tsinfer\ and \relate\ would seem more appropriate.
% [More]

% % Is it practical to estimate full ARGs?
% % [Not particularly \citep{shipilina2022origin} - but maybe ARGneedle
% % has done this for UKB? Need to check]

% % Do downstream applications actually make use of such full ARGs?
% Downstream applications using estimated ARGs to date rarely
% make use of such complete recombination information.
% The \relate\ selection statistic~\citep{speidel2019method}
% computes a $p$-value for selection [FIXME] for each
% local tree independently by comparing the rates of coalescence
% at various time-points. For cases like strong positive selection,
% where we expect there to be a clear signal in the local tree,
% such tree-by-tree methods can be simple and powerful.
% In their method
% for estimating dispersal rates and the locations of genetic
% ancestors,
% \citep{osmond2021estimating} downsample trees along the genome
% so that they can be regarded as approximately independent.
% The SIA method for detecting selection~\citep{hejase2022deep}
% encodes local trees as a set of lineage counts at discrete
% time intervals, and uses these as feature for a
% type of machine learning algorithm
% that takes ``temporal'' correlations into account. Thus,
% while SIA does use some of the information about local tree correlation,
% clearly much of the detail about recombination
% events in a ``full'' ARG is lost.
% The \texttt{tsdate} algorithm (and related naive estimator
% for ancestral location) uses much more between-tree
% information~\citep{wohns2022unified}, explicitly using the gARG
% encoding to reason about ancestral nodes.
% A major strength of this approach is that it can be used on
% any gARG, ranging from estimates containing relatively little
% information about recombination (e.g., as estimated by \relate)
% to ``full'' ARGs containing complete recombination information
% (e.g., as estimated by \kwarg), and will fully utilise
% any between-tree correlation information that is present.

\section*{Computational efficiency and data interchange}
ARG inference methods have been successfully applied to vast datasets,
including 500,000 genotyped humans~\citep{kelleher2019inferring,zhang2023biobank}
from UK Biobank~\citep{bycroft2018genome}
and over a million SARS-Cov-2 whole genomes~\citep{zhan2023towards}.
Even larger datasets are
available~\citep[e.g.][]{halldorsson2022sequences} % TODO more
and it seems inevitable that ARGs will be inferred for them.
The efficiency with which we can store and analyse these inferred
ARGs is therefore of great importantance, and a key factor
in the ultimate success of downstream ARG-based methods.
Another key factor in the how important ARG-based methods become
across the wide range of predicted application
areas [CITE commentary pieces, tsinfer, ARGneedle any others?]
is how well inference methods and downstream analysis applications
interoperate. [Genetics doesn't have a good record here, let's
try not to make the same mistakes again.]

The gARG encoding discussed and concretely defined in this manuscript
leads to highly efficient storage and processing of ARG data,
and has already been in use for several years.
The succinct tree sequence data structure (usually known as a ``tree sequence''
for brevity, but see below for confusion around this point)
is a concrete encoding of a gARG, as discussed in this article.
It was originally developed as part of the \texttt{msprime}
simulator~\cite{kelleher2016efficient} and subsequently been
extended and applied to forward-time
simulations~\citep{kelleher2018efficient,haller2018tree},
calculation of population genetics statistics~\citep{ralph2020efficiently}.
The succinct tree sequence encoding extends the basic definition
of a gARG provided here by stipulating a
simple tabular representation and nodes and edges,
and also defining a concise and lossless representation of
sequence variation using the ``site'' and  ``mutation'' tables.

The \texttt{tskit} library~\citep{tskit2023tskit} is a liberally
licensed open source toolkit that provides a comprehensive suite
of tools for working with gARGs (encoded as a succinct tree sequence).
Based on core functionality written
in C, it provides interfaces in C, Python and Rust.
Tskit is mature software, widely used in population genetics, and
has been incorporated into several downstream
applications~\citep[e.g.,][]{haller2019slim,speidel2019method,
adrion2020community,
terasaki2021geonomics,
baumdicker2021efficient,
fan2022genealogical,korfmann2022weak,
mahmoudi2022bayesian,petr2022slendr,rasmussen2022espalier,
zhang2023biobank}.

In retrospect, the choice of terminology around the succinct
tree sequence data structure was unfortunate, and has led to
significant confusion. [Discuss, and finish on a positive note]

@@

Reconstructing a most parsimonious ARG for a given data set is NP-hard
\citep{wang2001perfect}, so parsimony-based methods resort to heuristics and
are limited to analysing at most hundreds of sequences. Hence, a number of
methods aim to balance computational efficiency with reconstruction of
``reasonable", rather than parsimonious ARGs \citep{minichiello2006mapping,
parida2008estimating, kelleher2019inferring,  speidel2019method,
zhang2023biobank}. The latter three methods, as well as the parsimony-based
method of \citet{rasmussen2022espalier}, can be used on megabase scale data and
hundreds of thousands of samples under human-like parameters.
